Layer dependence as a measure of local dependence

### 2015-04-23 ld10 - abstract.tex
second submission

@@

\section{Local dependence and layer dependence}\label{sreview}

Dependence between two random variables generally varies with percentile. For example extreme movements in  stock markets are likely to be highly related  whereas minor fluctuations may be relatively independent.  Catastrophes create significant insurance losses for several classes of business at the same time, while attritional losses are typically weakly dependent.

Local dependence measures aim to capture the dependence structure of a bivariate distribution. This contrasts with  measures of overall dependence such as Pearson correlation, Spearman's $\rho$ and Kendall's $\tau$ \citep{embrechts2002correlation}. Local dependence measures include the univariate tail concentration \citep{venter2002tails}, correlation curve \citep{bjerve1993correlation}, and bivariate measures by \cite{bairamov2003new}, \cite{jones1996local} and \cite{holland1987dependence}.

This paper introduces, illustrates and analyzes an alternate local dependence measure called  ``layer dependence." Layer dependence is the covariance between a random variable and a single ``layer" of another. Layer dependence is also the ``gap" between upper and lower conditional tail expectations. Layer dependence is calculated  from the copula underlying the joint distribution.  Hence of interest is rank dependence rather than dependence between random variables in their original scale: the latter is often distorted by marginal distributions.

Layer dependence satisfies ``coherence" properties similar to Spearman's $\rho$: it is between $-1$ and $1$, constant and equal to $-1$, $0$ and $1$ for countermonotonic, independent and comonotonic random variables, sign switching when the ranking order reverses, and taking on higher values when dependence is stronger. Taking a weighted average of layer dependence values across the joint distribution yields Spearman's $\rho$ and alternative coherent measures of overall dependence.

Layer dependence provides a more appropriate and accurate measure of local dependence compared to existing measures. Higher dispersion between scatter points from the $45^\circ$ line reduces layer dependence and vice versa. For a Gumbel copula exhibiting upper tail dependence, layer dependence starts from a low value and increases to $1$ while the opposite applies to a Clayton copula with lower tail dependence.

Calculating layer dependence at the first instance extracts essential and interpretable information -- the dependence structure. This is often more informative than positing parametric copula, as the implication of its parametric form and parameters on the dependence structure is indirect. Computed layer dependence facilitates the selection and fitting of an appropriate copula.

Remaining sections are as follows. Section \aref{sintroduction} discusses the concepts leading to definition of layer dependence. Section \aref{sldcurve} demonstrates how layer dependence extracts the dependence structure from common copulas. Section \aref{sdecompose} explains the behaviour of layer dependence by decomposing it into a negative function of discordance and dispersion. Section \aref{scoherence} describes coherence properties of layer dependence. Links to existing tail dependence measures are highlighted in  \sref{sliterature}. Further properties of layer dependence are described in  \sref{sproperties}. Section \aref{saltoverall} forms alternate coherent measures of overall dependence apart from Spearman's $\rho$, using weighted averages of layer dependence. Section \aref{sfitting} discusses how layer dependence can be applied to copula modeling. Section \aref{sconclusion} concludes.

\begin{comment}
The literature suggests several approaches on fitting a copula to past data. If marginal distributions are unknown, a parametric approach typically involves selecting parametric copula and marginal distributions, and estimating parameters by maximising joint likelihood \citep{denuitactuarial}. A semi-parametric approach replaces marginal distributions in joint likelihood with empirical values \citep{oakes1989bivariate}. The choice of parametric copula may be restricted to a specific class of copulas. \cite{genest1993statistical} suggests an approach to select the generator function of Archimedean copulas \citep{mcneil2005qrm}. Alternatively a visual assessment of data may suggest an appropriate parametric copula with similar dependence structure, for example the Gumbel copula if upper tail dependence is present and Clayton copula if lower tail dependence is present. At the other extreme of copula fitting is to use the empirical copula, when the volume of past data is sufficiently large. \citep{czado2010pair} discusses a semi-parametric approach for multivariate copulas, based on vine copulas.
\end{comment}

\section{Layer dependence -- motivation and definition}\label{sintroduction}

A familiar construct in the study of bivariate dependence is Spearman's correlation \citep{embrechts2002correlation} defined as the linear correlation between ranks of two random variables. Rank dependence avoids distortion arising from marginal distributions as with for example Pearson's correlation \citep{mcneil2005qrm} measuring the degree of linear relationship between random variables in their original scale. Spearman's correlation can also be applied to estimate copula parameters using the method of moments (\cite{kojadinovic2010comparison}, \cite{bouye2000copulas}). However Spearman's correlation suffers from shortcomings and, as an aggregate measure, is inappropriate for assessing local dependence when dependence varies across the joint distribution including the tails.

Another familiar construct, in reinsurance, is a loss layer \citep{wang1995insurance}.  For example the 95\%--96\% layer of a random continuous loss $x$ is the portion of $x$ between its 95th and 96th percentile
$$
\min\{(x-x_{0.95})^+,x_{0.96}-x_{0.95}\}\ ,
$$
where $()^+$ indicates the positive part of the expression inside the brackets and subscripts indicate the percentile. Layers also arise in the context of derivative payouts and debt tranches in collaterised debt obligations \citep{mandel2012role}.  With rank dependence, the above quantities the 95\%--96\% layer reduces to
\begin{equation}\label{eg}
\min\{(u-u_{0.95})^+,u_{0.96}-u_{0.95}\} = \min\{(u-\alpha)^+,\de\alpha\}\ ,
\end{equation}
where $u$ is the percentile rank  and in this example $\alpha=0.95$ and $\de \alpha=0.01$.  Note $u_\alpha=\alpha$.

\subsection{Percentile layer decomposition}

The final expression in \eref{eg} can be written and approximated by, using the familiar infinitesimal notation,
\begin{equation}\label{ld}
-\de (u-\alpha)^+ \approx I_\alpha(u)\de \alpha  \cq I_\alpha(u) \equiv \left\{\begin{array}{cc}0\ , & u\le\alpha\ ,\\ 1\ ,& u>\alpha\ .\end{array}\right.
\end{equation}
The left hand side expression in \eref{ld} is called the $\alpha$--layer of $u$. The approximation becomes exact  as $\de\alpha \rightarrow 0$. The $\alpha$--layer of $u$ is an infinitesimally small increment $\de\alpha$ if $u$ exceeds $\alpha$ and is zero otherwise. Hence the $\alpha$--layer captures the movement in $u$ at $\alpha$ and ignores movements elsewhere.  \fref{flayer} illustrates the $\alpha$--layer of $u$.
\begin{figure}
  \begin{center}
    \begin{tabular}{c}
      \resizebox{100mm}{!}{\includegraphics{layer.eps}}
    \end{tabular}
    \caption{Illustration of $\alpha$-layer of $u$, written as $I_\alpha(u)\de \alpha$. The $\alpha$-layer of $u$ is only sensitive to movements in $u$ at $\alpha$ and ignores other movements. (*** Maybe u axis should stop at 1 *** }
    \label{flayer}
  \end{center}
\end{figure}

As often exploited  in reinsurance, any non-negative random loss or variable can be thought of as a sum of layers:
\begin{equation}\label{decompose}
u=\int_0^u 1\de \alpha= \int_0^1 I_\alpha(u)\de\alpha    \;.
\end{equation}
Hence $u$ is formed from infinitely many $\alpha$--layers, each capturing the movement of $u$ at a different $\alpha$.

\subsection{Constructing layer dependence}

Spearman's correlation is the correlation and also the standardized covariance between two percentile rank random variables $v$ and $u$:
\begin{equation}\label{rho}
\rho \equiv \cor(u,v) = \frac{\cov(v,u)}{\sqrt{\cov(v,v)\cov(u,u)}}=\frac{\cov(v,u)}{\cov(u,u)} = 12\cov(v,u)\ ,
\end{equation}
where $\cor$ and $\cov$ indicate covariance and correlation, respectively. In this paper assume the copula of $u$ and $v$ is exchangeable \citep{nelson1999ic}.

Using the decomposition of $u$ in \eref{decompose},
\begin{equation}\label{spearman}
\cov(v,u)=\cov\left\{v,\int_0^1I_\alpha(u)\de \alpha \right\}= \int_0^1 \cov\{v,I_\alpha(u)\} \de \alpha \; .
\end{equation}
Hence the covariance $\cov(v,u)$ can be thought of as the sum  into infinitely many covariances $\cov\{v,I_\alpha(u)\}$ for $0\le\alpha\le 1$. Each covariance in the sum measures the dependence between $v$ and the $\alpha$--layer of $u$. In a reinsurance setting, this covariance measures dependence between a particular layer of a loss (in percentile rank terms) and another factor. Alternatively the covariance captures dependence between $v$ and movements in $u$ at $\alpha$.

 Similar to $\rho$, scaling the layer covariances  $\cov\{v,I_\alpha(u)\}$ with the same when $v=u$ leads to the definition of layer dependence
\begin{equation}\label{ellalpha}
\ell_\alpha \equiv \frac{\cov\{v,I_\alpha(u)\}}{\cov\{u,I_\alpha(u)\}}   \;,
\end{equation}
where the denominator
\begin{equation}\label{covui}
\cov\{u,I_\alpha(u)\}=\E[\{u-\E(u)\}I_\alpha(u)]=\int_\alpha^1 \left(u  - \frac{1}{2}\right)\de \alpha = \frac{\alpha(1-\alpha)}{2}\ .
\end{equation}
Combining \eref{rho}, \eref{spearman}, \eref{ellalpha}  and \eref{covui} yields
\begin{equation}\label{wtdavg}
\rho = 12\int_0^1 \ell_\alpha \cov\{u,I_\alpha(u)\} \de \alpha =  \int_0^1 \ell_\alpha 6\alpha(1-\alpha) \de \alpha \; .
\end{equation}

Hence $\rho$ is a weighted average of $\ell_\alpha$ for $0\le\alpha\le 1$ with weights $w_\alpha=6\alpha(1-\alpha)$ integrating to $1$, and $\ell_\alpha$ decomposes $\rho$ into local dependence values. Note $w_\alpha$ has minimum $0$ at $\alpha=0$ and $1$ and increases symmetrically to maximum at $\alpha=0.5$. Hence Spearman's correlation places little emphasis on the tails which may be undesirable in finance or insurance where tail dependence is critical. Modifying the weights $w_\alpha$ leads to alternate measures of overall dependence further discussed in \sref{saltoverall}.

Layer dependence summarises the dependence structure of a copula. Layer dependence provides additional information compared to Spearman's correlation: how dependence varies across the joint distribution. As  any summary measure, layer dependence can mislead but is less misleading than $\rho$ and other measures of overall dependence. Layer dependence is  a more meaningful characterisation of a copula compared to the parameters of copula families such as the Clayton or Gumbel \citep{mcneil2005qrm}. Properties of and arguments for using layer dependence are explored below.

\begin{comment}
(** Not sure this is needed **) This paper assumes exchangeability: $C(u,v)=C(v,u)$ where $C$ is the copula of $(u,v)$. Hence $u$ and $v$ can be interchanged in the above expressions and those remaining in this paper. Archimedean copulas are a well known class of exchangeable copulas \citep{mcneil2005qrm}.
\end{comment}

\subsection{Key properties of layer dependence}

\begin{comment}
Similar to correlation, $0\le\ell_\alpha\le 1$ with $\ell_\alpha=-1$, $0$ and $1$ if $u$ and $v$ are countermonotonic, independent and comonotonic, respectively. In addition $\ell_\alpha$ increases with the correlation order of $(u,v)$. Replacing $u$ or $v$ with their complement leads to straightforward changes in layer dependence.

Layer dependence $\ell_\alpha$ satisfies  coherence properties for all $0\le\alpha\le 1$. These properties are similar to those of Spearman's correlation and are formalised in \sref{scoherence}.
\end{comment}

In \sref{sldcurve} and \sref{sdecompose} it is illustrated and shown  that  $\ell_\alpha=1$  occurs if and only if $I_\alpha(u)$ and $I_\alpha(v)$ are both simultaneously 1 or 0: that is $u$ and $v$ are both  either greater than $\alpha$ or less than $\alpha$.  Additionally $\ell_\alpha=1$ for $c\le\alpha\le d$ implies $u=v$ over $c\le u,v\le d$.

Layer dependence $\ell_\alpha$ satisfies the following coherence properties for all $0\le\alpha\le 1$. These properties are shared with Spearman's correlation and are formalised in \sref{scoherence}. Similar to correlation, $0\le\ell_\alpha\le 1$, and $\ell_\alpha=-1$, $0$ and $1$ if $u$ and $v$ are countermonotonic, independent and comonotonic, respectively. In addition $\ell_\alpha$ increases with the correlation order of $(u,v)$. Replacing $u$ or $v$ with their complement leads to straightforward changes in layer dependence.

The following is an alternative expression for $\ell_\alpha$:
\begin{equation}\label{gapexp}
\ell_\alpha = \frac{\E(v|u>\alpha)-\E(v|u\leq \alpha)}{\E(u|u>\alpha)-\E(u|u\leq \alpha)}
=2 \left\{\E(v|u>\alpha)-\E(v|u\leq \alpha)\right\}\;.
\end{equation}
Proofs are in an appendix. The middle expression in \eref{gapexp} is the expected change in $v$ relative to the expected change in $u$ when $u$ crosses $\alpha$. The latter is $0.5$ for all $\alpha$, yielding the final expression in \eref{gapexp}. Hence large $\ell_\alpha$ implies $v$ is sensitive to movements in $u$ across $\alpha$, indicating strong dependence between $v$ and $u$ at $\alpha$. When $\ell_\alpha=0$, $v$ is unchanged on average when $u$ crosses $\alpha$.

\begin{comment}
  Further alternative expressions are
$\ell_\alpha = (\tau_\alpha-1)/\alpha$ and $\tau_\alpha= (1+\alpha\ell_\alpha)/2$
where  $\tau_\alpha\equiv \E(v|u>\alpha)$.  These  alternative expressions for $\ell_\alpha$ follow from
$$
\cov\{v,(u>\alpha)\} =\frac{(2\tau_\alpha-1)(1-\alpha)}{2}
\cq
\cov\{u,(u>\alpha)\} =\frac{\alpha(1-\alpha)}{2}\ .
$$

Further
$$
\cor\{v,(u>\alpha)\}=\frac{\cov\{v,(u>\alpha)\}}{\sqrt{\alpha(1-\alpha)/12}}  = \ell_\alpha \sqrt{3\alpha(1-\alpha)}\ .
$$
\end{comment}

\section{Illustration of layer dependence curves for various copulas}\label{sldcurve}

The nine panels in \fref{fillustration} display $(u,v)$ scatterplots of well known exchangeable copulas. Less standard copulas are also shown for illustration. All copulas are calibrated to have equal Spearman's correlation $\rho=0.6$. Layer dependence curves $\ell_\alpha$ for all $0\le\alpha\le 1$ are plotted against $\alpha$ on $(u,v)$ scatterplots to demonstrate the link between the scatter and $\ell_\alpha$.

The scatterplots in \fref{fillustration}  emphasize that copulas with the same overall dependence can exhibit a variety of local dependence structures and these structures are captured by layer dependence curves. Given $\alpha$, $\ell_\alpha$ is larger if scatter points are more clustered around $(\alpha,\alpha)$ and vice versa. In particular $\ell_\alpha$ increases to one in the tails of copulas exhibiting strong tail dependence. Hence $\ell_\alpha$ tracks the clustering of scatter points across the $45^\circ$ line. This  is formalised in \sref{sdecompose}.

\begin{figure}
  \begin{center}
    \begin{tabular}{ccc}
      \resizebox{40mm}{!}{\includegraphics{normal.eps}}
            \resizebox{40mm}{!}{\includegraphics{student.eps}}
      \resizebox{40mm}{!}{\includegraphics{structural2.eps}} \\
      \resizebox{40mm}{!}{\includegraphics{gumbel.eps}}
            \resizebox{40mm}{!}{\includegraphics{structural1.eps}}
      \resizebox{40mm}{!}{\includegraphics{clayton.eps}} \\
            \resizebox{40mm}{!}{\includegraphics{frank.eps}}
            \resizebox{40mm}{!}{\includegraphics{structural3.eps}}
      \resizebox{40mm}{!}{\includegraphics{structural4.eps}} \\
    \end{tabular}
    \caption{Copulas all with the same $\rho=0.6$ but different layer dependence curves $\ell_\alpha$ over $0\leq\alpha\le 1$ (in red). The top 3 copulas display relatively strong tail dependence in both lower and upper tail.  The first two copulas in the middle three have relatively strong upper tail dependence and while the third has a high degree of lower tail dependence. The bottom three copulas have relatively high local dependence in the middle of the distribution and less correlation in the tails. Overall the panels illustrate how a given overall level of $\rho$ can mask a range of dependence structures and that layer dependence is an appropriate tool for assessing local dependence. The top left panel displays a Gaussian copula, followed by a Student's $t$ copula on the right. The leftmost panel in the middle three is a Gumbel copula, and the rightmost is a Clayton copula. The bottom left panel is a Frank copula. Remaining copulas are constructed from a factor model and are included to illustrate how layer dependence captures local dependence.}
    \label{fillustration}
  \end{center}
\end{figure}

\section{Discordance and dispersion}\label{sdecompose}

Layer dependence is intimately connected to the discordance and dispersion.
Assuming the copula $C$ of $(u,v)$ is exchangeable then
\begin{equation}\label{decomposition}
\ell_\alpha = 1-2(1+\gamma_\alpha) \delta_\alpha \;,
\end{equation}
where
$$
\gamma_\alpha \equiv \cor\{1-I_\alpha(u),I_\alpha(v)\} =\cor\{I_\alpha(u),1-I_\alpha(v)\}=\frac{\alpha^2-C(\alpha,\alpha)}{\alpha(1-\alpha)}\;,
$$
and
$$
\delta_\alpha \equiv\E\left\{(|u-v|)|(u-\alpha)(v-\alpha)<0\right\}  \; .
$$
A proof of \eref{decomposition} is  in the appendix. Equation \eref{decomposition}, as shown below, explains the behaviour of layer dependence curves in \fref{fillustration} where $\ell_\alpha$ for any $\alpha$ increases as the scatter become more tightly concentrated  around the point $(\alpha,\alpha)$, that is lower discordance and dispersion relative to the same point.

The correlation $-1\leq\gamma_\alpha\leq 1$ measures the tendency for $(u,v)$ to be discordant at $\alpha$: either $u$ or $v$ is above $\alpha$ and the other is below $\alpha$. The expectation $0\leq \delta_\alpha\leq 1$ measures the average dispersion between discordant $u$ and $v$ at $\alpha$, noting $(u-\alpha)(v-\alpha)<0$ is equivalent to $u>\alpha$ and $v\le\alpha$ or $u\le\alpha$ and $v>\alpha$.

\fref{finterpretation} illustrates \eref{decomposition} using two copulas specifically constructed such that, for $\alpha=0.75$, $\ell_\alpha=1$ for one and $\ell_\alpha=0.86$ for the other. For the copula where $\ell_\alpha=1$, there is no discordance or dispersion at $\alpha$. This observation is confirmed by setting $\ell_\alpha=1$ in \eref{decomposition} yielding $\gamma_\alpha=-1$ or $\delta_\alpha=0$. In this case $u$ and $v$ are perfectly dependent at $\alpha$, and are simultaneously below or above $\alpha$. It is also straightforward to extend this result such that if $\ell_\alpha=1$ over $c\le\alpha\le d$ then $u=v$ over $c\le u,v\le d$. As $\ell_\alpha$ decreases from $1$, as for the second copula, the extent of discordance $\gamma_\alpha$ and dispersion $\delta_\alpha$ increases. Again this result can be confirmed from \eref{decomposition} noting $\ell_\alpha$ is negatively related to $\gamma_\alpha$ and $\delta_\alpha$.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \resizebox{60mm}{!}{\includegraphics{perfect.eps}}
      \resizebox{60mm}{!}{\includegraphics{imperf.eps}} \\
    \end{tabular}
    \caption{The left and right panel show copulas where $\ell_{0.75}=1$ and $\ell_{0.75}=0.86$, respectively. In the left panel, $\gamma_{0.75}=-1$ (no discordant points) and $\delta_{0.75}=0$ (zero dispersion). In the right panel, $\gamma_{0.75}=-0.65$ (some discordant points) and $\delta_{0.75}=0.21$ (some dispersion between discordant points). }
    \label{finterpretation}
  \end{center}
\end{figure}

Applying \eref{decomposition} and generalising the illustration in \fref{finterpretation} explains the behaviour of layer dependence curves in \fref{fillustration}. Layer dependence $\ell_\alpha$ is larger if scatter points are more clustered around $(\alpha,\alpha)$, that is, fewer discordant points at $\alpha$, and discordant points at $\alpha$ are closer to the $45^\circ$ degree line. The former indicates smaller $\gamma_\alpha$ and the latter indicates smaller $\delta_\alpha$. Opposite observations apply for small $\ell_\alpha$.

\begin{comment}

The probability of discordance is
$$
\E[\{(u-\alpha)(v-\alpha)<0\}]
=2\alpha\{(1-\alpha)(\gamma_\alpha+\alpha^2)-1\} \;.
$$

The following illustrates \eref{decomposition} using the income-education example introduced in \sref{sintroduction}. Given $\alpha$, classify individuals as ``concordant" or ``discordant." The former are those with income and education both below or above $\alpha$. The latter are remaining individuals. Scale the proportion of concordant individuals to yield $\gamma_\alpha^*$. Calculate average dispersion $\delta_\alpha$ as twice the average difference between income and education of discordant individuals. Combine $\gamma^*_\alpha$ and $\delta_\alpha$ using \eref{decomposition} to yield $\alpha$-layer dependence $\ell_\alpha$. Layer dependence decreases with the proportion of discordant individuals and their average income-education dispersion.

\fref{fdecomposition} illustrates the decomposition of percentile rank gap by separately showing the standardised concordance probability $\gamma_\alpha^*$ and dispersion $\delta_\alpha$. The same four copulas as in \fref{fillustration} are applied.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \resizebox{60mm}{!}{\includegraphics{normaldecom.eps}}
      \resizebox{60mm}{!}{\includegraphics{gumbeldecom.eps}} \\
      \resizebox{60mm}{!}{\includegraphics{claytondecom.eps}}
      \resizebox{60mm}{!}{\includegraphics{frankdecom.eps}} \\
    \end{tabular}
    \caption{Standardised concordance probability $\sgamma^*_\alpha$ (red line) and dispersion $\delta_\alpha$ (blue line) for a Gaussian copula (top left), Gumbel copula (top right), Clayton copula (bottom left) and Frank copula (bottom right). Percentile rank gap (dotted black line) is also shown.}
    \label{fdecomposition}
  \end{center}
\end{figure}

Manipulating the expression for $\ell_\alpha$ in \eref{relationship} yields
$$
\ell_\alpha = 1- \frac{\E\{d(u,v) z_\alpha\}}{\alpha(1-\alpha)}  \cq \ell = 1-6\int_0^1 \E\{d(u,v)z_\alpha\} \de \alpha \;.
$$
where the above result for $\ell$ applies $\ell=\int_0^1 w_\alpha \ell_\alpha \de \alpha$, $w_\alpha\equiv 6\alpha(1-\alpha)$. Hence $\rho_S$ is negatively related to the average dispersion between discordant $(u,v)$ over all $\alpha$.

========== Haven't done much below here =============

\subsection{Income-education interpretation}

The following interprets layer dependence using an income-education example. Consider education $(u)$ and income $(v)$ rankings of individuals in a population, with $0$ and $1$ representing lowest and highest, respectively. Given $\alpha$, divide the population into two segments based on education: $u\leq \alpha$ and $u>\alpha$. Measure average income in each segment. From \eref{gapexp}, $\alpha$-layer dependence is twice the gap in average income between the two segments.

Suppose $\ell_\alpha$ is close to one for $\alpha=0.95$. This implies highly educated individuals (top $5\%$) are likely to be high income earners (top $5\%$), and vice versa. Few individuals have top $5\%$ education and income not in top $5\%$ or vice versa, and their average education-income difference is small. Hence there is strong income-education dependence at the $95$th percentile education layer. Dependence is perfect if $\ell_\alpha=1$. If $\ell_\alpha=0$ then average income is equal between individuals with education below and above top $5\%$, implying independence at the $95$th percentile education layer. The same argument applies to $\ell_\alpha$ for other values of $\alpha$.

This paper formulates layer dependence in terms of $u$ and $v$, in this case income and education ranks. Section \sref{soriginal} discusses an analogous definition of layer dependence between observed random variables $x$ and $y$, or original measurements of income and education.

\end{comment}

\section{Coherence properties of layer dependence}\label{scoherence}

Layer dependence $\ell_\alpha$ satisfies the following five ``coherence" properties. These properties are extensions of properties applying to Spearman's correlation.
\begin{itemize}

\item \textbf{Bounds}: Layer dependence lies between $-1$ and $1$: $-1 \le\ell_\alpha \le 1$ for all $\alpha$.
Hence layer dependence is bounded in the same way as $\rho$.

\item \textbf{Perfect dependence}: Constant layer dependence of $-1$ or $1$ are equivalent to countermonotonicity and comonotonicity, respectively. Thus $\ell_\alpha=-1$ for all $\alpha$ if and only if  $v=1-u$ while $\ell_\alpha=1$ for all $\alpha$ if and only if  $v=u$.

\item \textbf{Independence}: If $u$ and $v$ are independent then $\ell_\alpha=0$ for all $\alpha$. The converse is not true -- zero layer dependence does not imply independence as shown by the following counterexample. Assume $v=u$ and $v=1-u$ with equal probability. Then $\E(v|u=\alpha)=0.5$ for all $0\leq \alpha\leq 1$ implying $\E(v|u>\alpha)=\E(v|u\leq\alpha)=0.5$. Hence $\ell_\alpha=0$ from \eref{gapexp}. However $u$ and $v$ are not independent.

\item \textbf{Symmetry}: Replacing $v$ with $1-v$ yields layer dependence curve $-\ell_\alpha$. Doing the same to $u$ (the random variable decomposed into layers) yields layer dependence curve $-\ell_{1-\alpha}$ hence a flip is performed about $\alpha=0.5$ in addition to a sign change. Replacing both $u$ and $v$ with their complements yields layer dependence curve $\ell_{1-\alpha}$.

\item \textbf{Ordering}: Higher correlation order \citep{dhaene2009correlation} leads to higher layer dependence. Specifically, consider bivariate uniform $(u^*,v^*)$ exceeding $(u,v)$ in correlation order: $C^*(a,b)\geq C(a,b)$ for all $0\leq a,b\leq 1$, where $C^*$ is the copula of $(u^*,v^*)$. Then
$\ell^*_\alpha \geq \ell_\alpha$,   $0\leq\alpha\leq 1$ where $\ell_\alpha^*$ denotes the $\alpha$--layer dependence of $(u^*,v^*)$. Hence greater dependence leads to a higher layer dependence curve.

\end{itemize}
Independence and symmetry properties follow from the definition of layer dependence in \eref{ellalpha}. From \eref{gapexp}, constant layer dependence of one implies $\E(v|u>\alpha)=(\alpha+1)/2$ and $\E(v|u=\alpha)=\alpha$, for all $0\leq\alpha\leq 1$, hence $v=u$. Similarly constant layer dependence of minus one implies $v=1-u$. The ordering property holds since higher correlation order implies larger covariances \citep{dhaene2009correlation}. Prove the bounds property by combining ordering and perfect dependence properties, and noting countermonotonicity and comonotonicity represent minimum and maximum correlation order, respectively. Detailed proofs are shown in the appendix.

Most of the layer dependence coherence properties can be expressed using copulas shown in Table \aref{table1}.

\begin{table}[h]
  \begin{center}
\begin{tabular}{c|c|c }
 Copula & Description & LD \\
 \hline
 $uv$ & Independent $u$ and $v$ & 0 \\
 $\min(u,v)$ & Comonotonic $u$ and $v$ & 1 \\
 $\max(u+v-1,0)$ & Countermonotonic $u$ and $v$ & $-1$\\
 $u-C(u,1-v)$ & Replace $v$ with $1-v$ & $-\ell_\alpha$ \\
 $v-C(1-u,v)$ & Replace $u$ with $1-u$ & $-\ell_{1-\alpha}$ \\
 $u+v-1+C(1-u,1-v)$ & Replace $v$ and $u$ with complements & $\ell_{1-\alpha}$\\
 \hline
\end{tabular}
\caption{Layer dependence (LD) for different copulas and transformations.}\label{table1}
  \end{center}
\end{table}

\begin{comment}

\subsection{Copula comparison using percentile gap}

The following compares copulas based on their percentile gap values. The comparison yields a goodness-of-fit test when modeling past data using copulas. \fref{comparison} plots upper conditional tail expectations $\E(v|u>\alpha)=0.5(\alpha\ell_\alpha+1)$ of the copulas in \fref{fillustration} against the Gaussian copula. Given $\alpha$, points above the $45^{o}$ line indicates the copula has percentile gap exceeding that of the Gaussian copula. Vice versa for points below the $45^{o}$ line.

 we can easily simulate $\tau_\alpha$ paths as $(1/2)+\e^{(1-\alpha)}+\eps_\alpha$.

\begin{figure}
  \begin{center}
    \begin{tabular}{ccc}
      \resizebox{40mm}{!}{\includegraphics{normal.eps}}
            \resizebox{40mm}{!}{\includegraphics{studentvs.eps}}
      \resizebox{40mm}{!}{\includegraphics{structural2vs.eps}} \\
      \resizebox{40mm}{!}{\includegraphics{gumbelvs.eps}}
            \resizebox{40mm}{!}{\includegraphics{structural1vs.eps}}
      \resizebox{40mm}{!}{\includegraphics{claytonvs.eps}} \\
            \resizebox{40mm}{!}{\includegraphics{frankvs.eps}}
            \resizebox{40mm}{!}{\includegraphics{structural3vs.eps}}
      \resizebox{40mm}{!}{\includegraphics{structural4vs.eps}} \\
    \end{tabular}
    \caption{Plot of upper conditional tail expectations $\E(v|u>\alpha)$ for given copulas against the same for the Gaussian copula.}
    \label{comparison}
  \end{center}
\end{figure}

\fref{fillustration1} expands \fref{fillustration} by graphing layer dependence curves for several parameters of Gaussian, Gumbel, Clayton and Frank copulas. Layer dependence curves summarise essential information (dependence structure) underlying each copula:
\begin{itemize}

\item Gaussian copula: symmetric U-shaped dependence with stronger dependence for larger $\ell$.

\item Gumbel copula: asymmetric U-shaped dependence with stronger dependence for larger $\theta$. Perfect upper tail dependence is always present.

\item Clayton copula: linearly declining dependence with stronger dependence for larger $\theta$. Perfect lower tail dependence is always present.

\item Frank copula: relatively stable dependence with stronger dependence for larger $\theta$.

\end{itemize}
Computing the layer dependence curve is an insightful and important exercise, particularly when the parametric form of a given copula is difficult to interpret, or when the dependence structure underlying an empirical copula is not apparent. Overall dependence measures such as  $\rho_S$ or Kendall's $\tau$ provide information on overall dependence rather than ``local dependence." Section \aref{sliterature} compares layer dependence with existing local dependence measures. It is shown that layer dependence provides a more accurate measure of local dependence. In addition layer dependence possesses coherence properties, further discussed in \sref{scoherence}.

\begin{figure}
  \begin{center}
    \begin{tabular}{cc}
      \resizebox{60mm}{!}{\includegraphics{gaussianmul.eps}}
      \resizebox{60mm}{!}{\includegraphics{gumbelmul.eps}} \\
      \resizebox{60mm}{!}{\includegraphics{claytonmul.eps}}
      \resizebox{60mm}{!}{\includegraphics{frankmul.eps}} \\
    \end{tabular}
    \caption{layer dependence curves for various parameters of a Gaussian copula (top left), Gumbel copula (top right), Clayton copula (bottom left) and Frank copula (bottom right). Parameter values are shown in legends.}
    \label{fillustration1}
  \end{center}
\end{figure}

\subsection{``Locality" of layer dependence}

layer dependence $\ell_\alpha$ measures correlation between $u$ in the neighbourhood of $\alpha$ and $v$. This ``locality" property is not apparent from the definition of $\ell_\alpha$. The following is a proof. Define $\ell_{[a,b]}$ as the weighted average
$$
\ell_{[a,b]} \equiv \frac{\int_a^b \ell_\alpha w_\alpha \de \alpha}{\int_a^b w_\alpha \de\alpha}
=\frac{\cov(v,u_{[a,b]})}{\cov(u,u_{[a,b]})}
\cq u_{[a,b]}\equiv \int_a^b (u>\alpha)\de\alpha \;,
$$
where $w_\alpha=6\alpha(1-\alpha)$ and $u_{[a,b]}$ is the ``$[a,b]$ layer" of $u$. Hence the weighted average $\ell_{[a,b]}$ is the scaled correlation between $[a,b]$ layer of $u$ and $v$. Taking the limit $b\rightarrow a^+$ of $\ell_{[a,b]}$ and setting $a=\alpha$ yields $\ell_\alpha$. Rewriting $u_{[a,b]}$ yields
$$
u_{[a,b]}=(u>a)\int_a^{\min(u,b)} \de \alpha = (u>a)\left\{\min(u,b)-a \right\} \;.
$$
Hence the $[a,b]$ layer of $u$ is constant and equal to $0$ when $u\leq a$ and $b-a$ when $u>a$. When $a<u\leq b$, $u_{[a,b]}=u-a$ implying the $[a,b]$ layer of $u$ moves in line with $u$. Therefore the weighted average $\ell_{[a,b]}$ measures correlation between movements of $u$ in the interval $[a,b]$ and $v$. Taking the limit $b\rightarrow a^+$ implies $\ell_\alpha$ measures correlation between $u$ in the neighbourhood of $\alpha$ and $v$. Hence layer dependence measures ``local" dependence.

\subsection{Connection to first-order conditional expectations}

layer dependence in \eref{definition} transforms first order conditional tail expectations into measures of local dependence between $u$ and $v$. Suppose the regression curve of $v$ on $u$, or $\E(v|u=t)$ for all $0\leq t\leq 1$, is given. Then $\alpha$-layer dependence is the integral
$$
\ell_\alpha = 2\left\{\frac{1}{1-\alpha}\int_\alpha^1 \E(v|u=t) \de t - \frac{1}{\alpha}\int_0^\alpha \E(v|u=t) \de t \right\} \;.
$$
Given $\alpha$-layer dependence yields first order conditional expectations:
$$
\E(v|u>\alpha)=\ell_\alpha \E(u|u>\alpha) + 0.5(1-\ell_\alpha)  \;,
$$
$$
\E(v|u=\alpha)=\alpha\ell_\alpha +0.5(1-\ell_\alpha)-0.5\alpha(1-\alpha)\ell'_\alpha \;.
$$
Hence first order conditional expectations are a weighted average of corresponding expectations assuming comonotonicity and independence, with weights $\ell_\alpha$ and $1-\ell_\alpha$, respectively. The conditional expectation $\E(v|u=\alpha)$ requires an additional adjustment term $0.5\alpha(1-\alpha)\ell'_\alpha$ based on the derivative $\ell'_\alpha$.

\end{comment}

\section{Connections to measures of tail dependence}\label{sliterature}

Measures have been proposed to capture the degree of tail dependence. Tail dependence is dependence between extreme values of random variables, in this case values of $u$ and $v$ near 0 or 1. Strong tail dependence creates catastrophic events such as  multiple bank failures and  market crashes. Layer dependence is intimately connected to two existing tail dependence measures -- coefficient of tail dependence and tail concentration function. \cite{sweeting2013calculating} and \cite{durante2014copulas} further discuss tail dependence measures.

\subsection{Tail dependence}

>>>>

\makeatletter
\def\ps@pprintTitle{%
\let\@oddhead\@empty
\let\@evenhead\@empty
\def\@oddfoot{}%
\let\@evenfoot\@oddfoot}
\makeatother

@@
